The imprint of global magma oceans on exoplanet demographics

### 2022-12-22 src/tex/ms.tex
"magma ocean hypothesis" -> "runaway greenhouse hypothesis"
...and relatives

The main task is to determine -- for different survey configurations -- the confidence level with which a magma ocean signal can be detected in simulated exoplanet observations.
For this purpose we expanded the Bioverse framework~\citep{Bixel2020,Bixel2021}\footnote{Bioverse is actively maintained and documented open source software written in Python. Its latest version and documentation can be found at \url{https://github.com/danielapai/bioverse}.} to generate synthetic stellar and planetary samples into which we injected magma ocean signals, simulated observations of this sample, and computed Bayesian evidences in favor of the magma ocean hypothesis (see Fig.~\ref{fig:flowchart}).
\begin{figure*}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/flowchart.pdf}
        \caption{Workflow of our hypothesis testing with Bioverse. In the first block, a stellar sample is generated based on the Gaia nearby stars catalog~\citep{Smart2021}\todo[]{KEVIN pls verify}. The stars are then populated with planets based on Kepler demographics, and they may be assigned a magma ocean based on the model described in Sect.~\ref{sec:mo_model}. The planets' respective transit probabilities are computed. The second block simulates the exoplanet survey whereby selection effects and detection biases are introduced. Finally, the third block deals with testing a hypothesis based on the data from the simulated survey. By iterating through these steps, we compute the statistical power of testing the hypothesis given the assumed survey design.}
        \label{fig:flowchart}
    \end{centering}
\end{figure*}

@@

    Central to our procedure is to explore the detectability of population-level trends with future exoplanet surveys, and here the trend in question is the proposed radius discontinuity.
    To enable a quantitative assessment of this detectability, we injected the signal into the simulated planet sample before observing it with simulated surveys.
    While we search for the magma ocean pattern in demographic quantities such as average planet radii, the injected changes happen on the planetary level: We changed each planet's radius based on its individual set of properties and the associated predictions from steam atmosphere and water incorporation models.
    Relevant properties are a planet's mass $M$, its instellation $S$, and its bulk water inventory expressed as a water mass fraction $x_{H_2O}$.
    We consider three different cases:
    \begin{itemize}
   \item[A.] No radius change.
    This case serves as our null hypothesis, where planetary radii are not affected by the amount of irradiation the planets receive.
    It also applies outside the runaway greenhouse limit, where we assume a solid planetary surface and no steam atmosphere.
        A planet that lies \textit{within} the runaway greenhouse limit but that is already completely desiccated will have the same radius. \todo{Mention if this case is considered in the model or not (as of now, no).}
   \item[B.] Radius inflation due to a steam atmosphere.
    We assigned planets within the runaway greenhouse limit steam atmospheres, which increase atmospheric scale heights significantly.
   \item[C.] As B, but including radius decrease of inflated planets due to water in the melt.
    Due to greenhouse forcing, planetary surfaces below steam atmospheres are thought to be molten.
    Part of the water will then partition into the melt, removing it from the steam atmosphere, which in turn decreases the radius inflation.
    This effect is generally small compared to the radius inflation from a steam atmosphere but resulting radius changes typically reach measurable values.
    \end{itemize}

@@

\subsection{Hypothesis tests}
We now turn to assessing the simulated surveys in terms of their ability to detect the runaway greenhouse transition and their constraining power on parameters of this trend.
To do this, we rely on a Bayesian hypothesis testing approach where we quantify the evidence of a hypothesis over another based on the (simulated) data.
For our specific problem, this implies comparing evidences for a demographic imprint of the runaway greenhouse effect to its absence.
\subsubsection{Definition of the runaway greenhouse hypothesis}
\begin{note}
%    Bayesian model comparison requires that an alternative mode be specified against which the comparison is made.
\end{note}
    As a null hypothesis, we consider the case where the planetary radius distribution is independent of the instellation,
    \begin{equation}
        h_{\mathrm{null}}(\theta, S) = \theta,
    \end{equation}
    where $\theta$ is the set of parameters defining the radius distribution.
    We further define an alternative hypothesis that describes radius changes due to steam atmospheres and water in the melt.
    As motivated above, this hypothesis takes the form of a step function in instellation $S$, where the step occurs at the outer edge of the runaway greenhouse region.
    Our main observable shall be the average planet radius in the planet population inside and outside this threshold.
    The magma ocean hypothesis is then defined as
\begin{equation}\label{eq:mo_hypo}
    h_{\mathrm{MO}}(\theta, S) =
%        \begin{cases}
%        1 & \text{if } x \in \mathbb{Q}\\
%        0 & \text{if } x \in \mathbb{R}\setminus\mathbb{Q}
%    \end{cases}
        \begin{cases}
            h_{\mathrm{null}} & \text{for } S \leq S_\mathrm{thresh}\\
            \langle R_\mathrm{P}\rangle (f_\mathrm{steam}, f_\mathrm{water}) & \text{for } S > S_\mathrm{thresh}.
        \end{cases}
\end{equation}
    Here, $f_\mathrm{steam}$ and $f_\mathrm{water}$ are predictions from the steam atmosphere and water incorporation models, respectively.
    They are assumed to act additively on the planet radii and thus on their average $\langle R_\mathrm{P}\rangle $.
\begin{figure}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/HnullHmo.pdf}
        \caption{Planets above and below the runaway greenhouse threshold.
        \textit{Top}: Planet state as a function of orbital period. Planets with and without a runaway greenhouse climate mix and are not distinguishable in period space.
        \textit{Bottom}: Radii of synthetic planets with injected radius deviation as a function of instellation. Only the planets marked as transiting are observable. Above the runaway greenhouse threshold $S_\mathrm{thresh} =\SI{280}{\watt\per\meter\squared} $, radii are inflated by the amount indicated with gray lines. The sharp boundary at $S_\mathrm{thresh}$ causes a discontinuity in the average planet radius (blue). This magma ocean hypothesis can be tested against the null hypothesis $H_0$ (green), where average radii are independent of instellation.}
        \label{fig:HnullHmo}
    \end{centering}
\end{figure}
    Figure~\ref{fig:HnullHmo} shows the null and alternative hypotheses as a function of instellation.
    The value for the average radius \textit{outside} the runaway greenhouse regime was computed from our nominal observed sample, whereas the radius change in the magma ocean case compared to the null hypothesis is a prediction from the magma ocean model, evaluated on the same sample.

\begin{note}
    The only free parameter of the null hypothesis, which assumes the average radius to be independent of instellation, is the predicted mean radius $\langle R_\mathrm{P}\rangle $.
    The functional form of the magma ocean hypothesis is more complex: Besides the mean radius of planets outside the runaway greenhouse threshold $\langle R_\mathrm{P}\rangle_\mathrm{out}$, which is a nuisance parameter necessary to define the hypothesis, it relies on the threshold instellation for the ``step'' $S_\mathrm{thresh}$.
    Further, the parameter encoding the amount of radius change is the planetary water mass fraction~$x_{H_2O}$.
\end{note}

\begin{note}
The measured radii $R_\mathrm{P, i}$ cannot be directly used for the hypothesis tests as they include intrinsic scatter that is not caused by measurement errors.
$h_{\mathrm{MO}}$ and $h_{\mathrm{null}}$ should thus be tested against a statistical estimator that represents the population mean.
To avoid binning and the artificial patterns it may introduce, we chose to test our hypotheses against a simple moving average $SMA$ along the instellation axis with a window of size \windowsize\ centered around each measurement.
We further computed the uncertainty of this moving average by propagating the individual measurement errors and applying a rolling standard error of the mean. \todo[inline]{specify error propagation}

As our procedure involves random sampling of the model parameters $\theta$, we need to define the probability of obtaining a dataset given the model parameters, i.e., a likelihood function.
We assumed here that the individual moving averages $SMA_i$ are measured with a normally distributed uncertainty $\sigma_{SMA_i}$ and adopted a normal distribution
\begin{equation}
    \mathcal{L}(SMA \mid \boldsymbol{\theta})=\prod_{i}^{N} \frac{1}{\sqrt{2 \pi \sigma_{SMA_i}^{2}}} \exp \left(-\frac{\left(SMA_i - h\left(\boldsymbol{\theta}, S_i\right)\right)^{2}}{2 \sigma_{SMA_i}^{2}}\right).
\end{equation}
Here, $h\left(\boldsymbol{\theta}, S_i\right)$ corresponds to the functional form of the magma ocean hypothesis~(Eqn.~\ref{eq:mo_hypo}).
We tested this hypothesis against the null hypothesis $h_\mathrm{null} (\theta, S) = \theta$, which states that there is no relationship between the measured planet radii and received instellation.
\end{note}

\subsubsection{Bayesian model comparison}
We are now ready to assess the relative plausibility of $h_{\mathrm{MO}}$ and $h_{\mathrm{null}}$ given the synthetic data we have generated, assigning equal a~priori~ probabilities to these models.
This is done by comparing the Bayesian evidence $\mathcal{Z}$ of the models, which we estimate with the nested sampling~\citep{Skilling2004} algorithm \emph{dynesty}~\citep{Speagle2020}.
Our criterion to reject the null hypothesis is
\begin{equation}
\Delta \ln \mathcal{Z}  = \ln \mathcal{Z} - \ln \mathcal{Z}_\mathrm{null}  > 3.
\end{equation}
The chosen threshold roughly corresponds to the $p = 0.05$ threshold for statistical significance popular in frequentist approaches.\todo{CITATION}

@@

\section{Results}
\subsection{Signature of the magma ocean hypothesis}
\todo[inline]{show general detectability}

\begin{figure}[ht!]
%    \script{figurescript.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/optimistic_R-S.pdf}
        \caption{
        Detection of the runaway greenhouse threshold in the optimistic case.
        From simulated radius and instellation measurements of a large ($N = \Nplanets$) survey, we compute the moving average (blue confidence intervals) and fit the runaway greenhouse hypothesis to it (Eqn.~\ref{eq:mo_hypo}, random draws from the posterior in red).
            The pattern is detected with high significance.
        }
        \label{fig:optimistic_R-S}
    \end{centering}
\end{figure}
Figure~\ref{fig:optimistic_R-S} shows a prototypical detection of the runaway greenhouse threshold.
For the sake of clarity, we chose an optimistic scenario where the sample is large ($N = \Nplanets$) and the measurement uncertainties are small ($\sigma_{R} = \sigmaR, \sigma_{S} = \sigmaS$).
We assumed a water mass fraction of \wrr\ for each planet.
In this case, the pattern was detected with high significance ($\ln \mathcal{Z} \approx 100$).

@@

PLATO (PLAnetary Transits and Oscillation of stars) is an ESA mission designed to characterize terrestrial planets in the habitable zones of Sun-like stars, a goal to be achieved via long-term high-precision photometric monitoring of a sample of bright stars~\citep{Rauer2016}.
The PLATO team has released an estimate on the number of exoplanets that will be characterized in the course of the main survey mission (Matuszewski, priv. comm.).
In the FGK sample of the mission, a minimum of 500 Earth-sized (\SIrange{0.8}{1.25}{\rEarth}) planets are expected to be detected.
We base our assessment of PLATO's diagnostic power to test the magma ocean hypothesis on these estimates.

@@

\begin{note}
   The fidelity of a future detection or falsification of a magma ocean signature will depend on the significance with which the null hypothesis can be excluded.
   As a consequence, instrumentation and survey strategy aiming at testing the hypothesis should aim at maximizing the probability of a true positive detection given the existence of the effect.
   This is called the \textit{statistical power}: the probability of correctly rejecting a false null hypothesis.
   Before turning to mission design trades that influence the statistical power in Sect.~\ref{sec:mission-design-trades}, we explore its behavior as a function of various free parameters of the magma ocean model.
    ...
\begin{figure*}[ht!]
%    \script{figurescript.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/example_statistical_power.pdf}
        \caption{
        Statistical power of a magma ocean hypothesis test as a function of model parameters.
        }
        \label{fig:statistical_power}
    \end{centering}
\end{figure*}

@@

\subsection{Statistical power of different mission designs}\label{sec:statpower_missions}
\todo[inline]{show detectability of the magma ocean signature for different mission designs}

@@

The key findings of our work are as follows:
\begin{enumerate}
    \item The demographic imprint of the runaway greenhouse threshold should be detectable with transit measurements of a large enough sample and sufficient precision.
    \item We find that the planned PLATO transit survey will provide a sufficient sample to test the magma ocean hypothesis.
    \item The diagnostic power of transit missions in testing this hypothesis can be maximized by ...
    \item A detection of the feature will constrain the water inventory of rocky exoplanets, providing important context in the assessment of their habitability.
\end{enumerate}

>>>>

We determined -- for different survey configurations -- the confidence level with which the runaway greenhouse threshold can be detected in simulated exoplanet observations.
For this purpose we expanded the Bioverse framework~\citep{Bixel2020,Bixel2021}\footnote{Bioverse is actively maintained and documented open source software written in Python. Its latest version and documentation can be found at \url{https://github.com/danielapai/bioverse}.} to generate synthetic stellar and planetary samples into which we injected the runaway greenhouse signal, simulated observations of this sample, and computed Bayesian evidences in favor of the runaway greenhouse hypothesis (see diagram in Fig.~\ref{fig:flowchart}).
\begin{figure*}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/flowchart.pdf}
        \caption{Workflow of our hypothesis testing with Bioverse. In the first block, a stellar sample is generated based on the Gaia nearby stars catalog~\citep{Smart2021}\todo[]{KEVIN pls verify}. The stars are then populated with planets based on Kepler demographics, and they may be assigned a magma ocean based on the model described in Sect.~\ref{sec:mo_model}. The planets' respective transit probabilities are computed. The second block simulates the exoplanet survey whereby selection effects and detection biases are introduced. Finally, the third block deals with testing a hypothesis based on the data from the simulated survey. By iterating through these steps, we compute the statistical power of testing the hypothesis given the assumed survey design.}
        \label{fig:flowchart}
    \end{centering}
\end{figure*}

@@

    Central to our procedure is to explore the detectability of population-level trends with future exoplanet surveys, and here the trend in question is the proposed radius discontinuity.
    To enable a quantitative assessment of this detectability, we injected the signal into the simulated planet sample before observing it with simulated surveys.
    While we search for the runaway greenhouse pattern in demographic quantities such as average planet radii, the injected changes happen on the planetary level: We changed each planet's radius based on its individual set of properties and the associated predictions from steam atmosphere and water incorporation models.
    Relevant properties are a planet's mass $M$, its instellation $S$, and its bulk water inventory expressed as a water mass fraction $x_{H_2O}$.
    We consider three different cases:
    \begin{itemize}
   \item[A.] No radius change.
    This case serves as our null hypothesis, where planetary radii are not affected by the amount of irradiation the planets receive.
    It also applies outside the runaway greenhouse limit, where we assume a solid planetary surface and no steam atmosphere.
        A planet that lies \textit{within} the runaway greenhouse limit but that is already completely desiccated will have the same radius. \todo{Mention if this case is considered in the model or not (as of now, no).}
   \item[B.] Radius inflation due to a steam atmosphere.
    We assigned planets within the runaway greenhouse limit steam atmospheres, which increase atmospheric scale heights significantly.
   \item[C.] As B, but including radius decrease of inflated planets due to water in the melt.
    Due to greenhouse forcing, planetary surfaces below steam atmospheres are thought to be molten.
    Part of the water will then partition into the melt, removing it from the steam atmosphere, which in turn decreases the radius inflation.
    This effect is generally small compared to the radius inflation from a steam atmosphere but resulting radius changes typically reach measurable values.
    \end{itemize}

@@

\subsection{Hypothesis tests}
We now turn to assessing the simulated surveys in terms of their ability to detect the runaway greenhouse transition and their constraining power on parameters of this trend.
To do this, we rely on a Bayesian hypothesis testing approach where we quantify the evidence of a hypothesis over another based on the (simulated) data.
For our specific problem, this implies comparing evidences for a demographic imprint of the runaway greenhouse effect to its absence.
\subsubsection{Definition of the runaway greenhouse hypothesis}
\begin{note}
%    Bayesian model comparison requires that an alternative mode be specified against which the comparison is made.
\end{note}
    As a null hypothesis, we consider the case where the planetary radius distribution is independent of the instellation,
    \begin{equation}
        h_{\mathrm{null}}(\theta, S) = \theta,
    \end{equation}
    where $\theta$ is the set of parameters defining the radius distribution.
    We further define an alternative hypothesis that describes radius changes due to steam atmospheres and water in the melt.
    As motivated above, this hypothesis takes the form of a step function in instellation $S$, where the step occurs at the outer edge of the runaway greenhouse region.
    Our main observable shall be the average planet radius in the planet population inside and outside this threshold.
    The runaway greenhouse hypothesis is then defined as
\begin{equation}\label{eq:rgh_hypo}
    h_{\mathrm{rgh}}(\theta, S) =
        \begin{cases}
            h_{\mathrm{null}} & \text{for } S \leq S_\mathrm{thresh}\\
            \langle R_\mathrm{P}\rangle (f_\mathrm{steam}, f_\mathrm{water}) & \text{for } S > S_\mathrm{thresh}.
        \end{cases}
\end{equation}
    Here, $f_\mathrm{steam}$ and $f_\mathrm{water}$ are predictions from the steam atmosphere and water incorporation models, respectively.
    They are assumed to act additively on the planet radii and thus on their average $\langle R_\mathrm{P}\rangle $.
\begin{figure}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/HnullHmo.pdf}
        \caption{Planets above and below the runaway greenhouse threshold.
        \textit{Top}: Planet state as a function of orbital period. Planets with and without a runaway greenhouse climate mix and are not distinguishable in period space.
        \textit{Bottom}: Radii of synthetic planets with injected radius deviation as a function of instellation. Only the planets marked as transiting are observable. Above the runaway greenhouse threshold $S_\mathrm{thresh} =\SI{280}{\watt\per\meter\squared} $, radii are inflated by the amount indicated with gray lines. The sharp boundary at $S_\mathrm{thresh}$ causes a discontinuity in the average planet radius (blue). This runaway greenhouse hypothesis can be tested against the null hypothesis $h_\mathrm{null}$ (green), where average radii are independent of instellation.}
        \label{fig:HnullHmo}
    \end{centering}
\end{figure}
    Figure~\ref{fig:HnullHmo} shows the null and alternative hypotheses as a function of instellation.
    The value for the average radius \textit{outside} the runaway greenhouse regime was computed from our nominal observed sample, whereas the radius change in the magma ocean case compared to the null hypothesis is a prediction from the magma ocean model, evaluated on the same sample.

\begin{note}
    The only free parameter of the null hypothesis, which assumes the average radius to be independent of instellation, is the predicted mean radius $\langle R_\mathrm{P}\rangle $.
    The functional form of the runaway greenhouse hypothesis is more complex: Besides the mean radius of planets outside the threshold $\langle R_\mathrm{P}\rangle_\mathrm{out}$, which is a nuisance parameter necessary to define the hypothesis, it relies on the threshold instellation for the ``step'' $S_\mathrm{thresh}$.
    Further, the parameter encoding the amount of radius change is the planetary water mass fraction~$x_{H_2O}$.
\end{note}

\begin{note}
The measured radii $R_\mathrm{P, i}$ cannot be directly used for the hypothesis tests as they include intrinsic scatter that is not caused by measurement errors.
$h_{\mathrm{rgh}}$ and $h_{\mathrm{null}}$ should thus be tested against a statistical estimator that represents the population mean.
To avoid binning and the artificial patterns it may introduce, we chose to test our hypotheses against a simple moving average $SMA$ along the instellation axis with a window of size \windowsize\ centered around each measurement.
We further computed the uncertainty of this moving average by propagating the individual measurement errors and applying a rolling standard error of the mean. \todo[inline]{specify error propagation}

As our procedure involves random sampling of the model parameters $\theta$, we need to define the probability of obtaining a dataset given the model parameters, i.e., a likelihood function.
We assumed here that the individual moving averages $SMA_i$ are measured with a normally distributed uncertainty $\sigma_{SMA_i}$ and adopted a normal distribution
\begin{equation}
    \mathcal{L}(SMA \mid \boldsymbol{\theta})=\prod_{i}^{N} \frac{1}{\sqrt{2 \pi \sigma_{SMA_i}^{2}}} \exp \left(-\frac{\left(SMA_i - h\left(\boldsymbol{\theta}, S_i\right)\right)^{2}}{2 \sigma_{SMA_i}^{2}}\right).
\end{equation}
Here, $h\left(\boldsymbol{\theta}, S_i\right)$ corresponds to the functional form of the runaway greenhouse hypothesis~(Eqn.~\ref{eq:rgh_hypo}).
We tested this hypothesis against the null hypothesis $h_\mathrm{null} (\theta, S) = \theta$, which states that there is no relationship between the measured planet radii and received instellation.
\end{note}

\subsubsection{Bayesian model comparison}
We are now ready to assess the relative plausibility of $h_{\mathrm{rgh}}$ and $h_{\mathrm{null}}$ given the synthetic data we have generated, assigning equal a~priori~ probabilities to these models.
This is done by comparing the Bayesian evidence $\mathcal{Z}$ of the models, which we estimate with the nested sampling~\citep{Skilling2004} algorithm \emph{dynesty}~\citep{Speagle2020}.
Our criterion to reject the null hypothesis is
\begin{equation}
\Delta \ln \mathcal{Z}  = \ln \mathcal{Z} - \ln \mathcal{Z}_\mathrm{null}  > 3.
\end{equation}
The chosen threshold roughly corresponds to the $p = 0.05$ threshold for statistical significance popular in frequentist approaches.\todo{CITATION}

@@

\section{Results}
\subsection{Signature of the runaway greenhouse hypothesis}
\todo[inline]{show general detectability}

\begin{figure}[ht!]
%    \script{figurescript.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/optimistic_R-S.pdf}
        \caption{
        Detection of the runaway greenhouse threshold in the optimistic case.
        From simulated radius and instellation measurements of a large ($N = \Nplanets$) survey, we compute the moving average (blue confidence intervals) and fit the runaway greenhouse hypothesis to it (Eqn.~\ref{eq:rgh_hypo}, random draws from the posterior in red).
            The pattern is detected with high significance.
        }
        \label{fig:optimistic_R-S}
    \end{centering}
\end{figure}
Figure~\ref{fig:optimistic_R-S} shows a prototypical detection of the runaway greenhouse threshold.
For the sake of clarity, we chose an optimistic scenario where the sample is large ($N = \Nplanets$) and the measurement uncertainties are small ($\sigma_{R} = \sigmaR, \sigma_{S} = \sigmaS$).
We assumed a water mass fraction of \wrr\ for each planet.
In this case, the pattern was detected with high significance ($\ln \mathcal{Z} \approx 100$).

@@

PLATO (PLAnetary Transits and Oscillation of stars) is an ESA mission designed to characterize terrestrial planets in the habitable zones of Sun-like stars, a goal to be achieved via long-term high-precision photometric monitoring of a sample of bright stars~\citep{Rauer2016}.
The PLATO team has released an estimate on the number of exoplanets that will be characterized in the course of the main survey mission (Matuszewski, priv. comm.).
In the FGK sample of the mission, a minimum of 500 Earth-sized (\SIrange{0.8}{1.25}{\rEarth}) planets are expected to be detected.
We base our assessment of PLATO's diagnostic power to test the runaway greenhouse hypothesis on these estimates.

@@

\begin{note}
   The fidelity of a future detection or falsification of a magma ocean signature will depend on the significance with which the null hypothesis can be excluded.
   As a consequence, instrumentation and survey strategy aiming at testing the hypothesis should aim at maximizing the probability of a true positive detection given the existence of the effect.
   This is called the \textit{statistical power}: the probability of correctly rejecting a false null hypothesis.
   Before turning to mission design trades that influence the statistical power in Sect.~\ref{sec:mission-design-trades}, we explore its behavior as a function of various free parameters of the magma ocean model.
    ...
\begin{figure*}[ht!]
%    \script{figurescript.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/example_statistical_power.pdf}
        \caption{
        Statistical power of the runaway greenhouse hypothesis test as a function of model parameters.
        }
        \label{fig:statistical_power}
    \end{centering}
\end{figure*}

@@

\subsection{Statistical power of different mission designs}\label{sec:statpower_missions}
\todo[inline]{show detectability of the runaway greenhouse signature for different mission designs}

@@

The key findings of our work are as follows:
\begin{enumerate}
    \item The demographic imprint of the runaway greenhouse threshold should be detectable with transit measurements of a large enough sample and sufficient precision.
    \item We find that the planned PLATO transit survey will provide a sufficient sample to test the runaway greenhouse hypothesis.
    \item The diagnostic power of transit missions in testing this hypothesis can be maximized by ...
    \item A detection of the feature will constrain the water inventory of rocky exoplanets, providing important context in the assessment of their habitability.
\end{enumerate}
